---
title: "Practicum 3 - Increased Incentives for Water Savings P4P Program"
author: "David Segan, Marie Bouffard, Ethan Hoffman"
date: "5/19/2021"
output: html_document
---


```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```


```{r load, echo=TRUE}
library(DeclareDesign)
library(truncnorm) #for truncated distribution
library(knitr)
library(ggplot2)
library(kableExtra)
```

##declare_population()

We are simulating a population of 1,000 buildings (our unit of analysis), each of which has a differenct level of water saved, measured in million gallons, and a different amount of variability in that water saved (the variable 'u'). 
```{r population, echo=TRUE}

set.seed(228)
population <- declare_population(
  # population of customers that use atleast 10 million gallons of water (minimum needed to be saved to be eligible for program)
  customers = add_level(N=1000, 
    # unit for water_saved is million gallons
    water_saved=rtruncnorm(n=N, a=10, b=Inf, mean=11, sd = 10),
    u=rnorm(n=N, mean=1.1, sd=0.1))
)

```

##Population descriptives

A truncated normal distribution is used because there are more buildings that save close to the 10 million gallon minimum program eligibility cutoff than there are that save much more.

```{r population-see, echo=TRUE, fig.height=5.5}
pop <- population()
hist(pop[,2], xlab="Baseline Water Saved (million gallons)", 
     main="Baseline", cex=24)
```

##declare_potential_outcomes()

The next step is to declare the full schedule of potential outcomes $Y(1)$ and $Y(0)$ under an assumption about the effect size of interest. We estimate that the customers in the higher incentive group save 1 million gallons more.

```{r po, echo=TRUE}

# treatment effect of higher incentive
te = 1 # million gallons

potential_outcomes <- 
  declare_potential_outcomes(
    Y_D_0=water_saved * u,
    Y_D_1=water_saved * u + te)

```


##Potential outcomes descriptives

```{r po-see, echo=TRUE}
po <- potential_outcomes(pop)
kable(po[1:5,], digits=1)
```


##declare_sampling()

We are using a sample size of 150 customers. 

```{r sample, echo=TRUE}
sampling <- declare_sampling(n=150)
sam <- sampling(po)
kable(sam[1:5,c(1:2,4:6)], row.names = FALSE,
      digits = 1)
```


##declare_assignment()

We randomly assignm 1/2 of sampled customers to the treatment.

```{r assign, echo=TRUE}
assigning <- declare_assignment(m = nrow(sam)/2,
                  assignment_variable="D")
assigned <- assigning(sam)
kable(assigned[1:5,c(1:2,4:5,7:8)], 
      digits = 1)

```


## Assessing balance

Review the balance in baseline water saved characteristics since random assignment has occured.

```{r violin, echo=FALSE, fig.height=6}
ggplot(data=assigned, aes(x=as.factor(D), y=water_saved)) +
geom_violin(aes(fill=as.factor(D), color=as.factor(D))) +
theme_minimal(base_size = 24) + xlab("Assignment")

```

##declare_reveal()

This step declares how the potential outcomes are revealed by the random assignment

```{r reveal, echo=TRUE}
revealing <- declare_reveal(assignment_variables=D)
```

##declare_estimand()

At this stage, we specify our target *estimand*, which is the quantity that we are trying to recover when estimating impact. Recall that we set this value to **1**

```{r estimand, echo=TRUE}
estimand <- declare_estimand(ATE = te)
estimand(po)
```


##declare_estimator()

Next, we declare the estimators we use for recovering the estimand. While there are many advanced estimators, we'll focus on the two core experimental estimators:
1. difference-in-means
2. difference-in-differences

```{r estimator, echo=TRUE}
dim <- declare_estimator(Y ~ D, estimand = estimand,  
          model =  difference_in_means, label = "DIM") #PH: Difference-in-means (Mean of treatment group - mean of control group)

did <- declare_estimator(Y - water_saved ~ D, 
                         estimand = estimand,  
          model =  difference_in_means, label = "DID") #PH: Difference-in-differences ([Mean of treatment group @ endline - Mean of treatment group @ baseline] - [Mean of control group @ endline - mean of control group @ baseline])
```


##declare_design()

This function brings all of the parts of the process together in a single design and allows for each part of the design to be simulated repeatedly.

```{r design, echo=TRUE}
design <- population + potential_outcomes + sampling +
          assigning + revealing + estimand + dim + did
```


##diagnose_design()

At this stage, we can calculate various features of the design that we have specified

```{r diagnosis, cache=TRUE}
diagnosis <- diagnose_design(design, sims=500)
diagnosis$diagnosands_df[,c(1,3,5,9,11)] %>%
  kable()
```


## Looking under the hood, DIM

```{r underhood-dim, height=6}
sim.out <- diagnosis$simulations
hist(sim.out$estimate[sim.out$estimator_label=="DIM"],
     main="Randomization Distribution",
     xlab="Estimates in Realized Experiments",
     xlim=c(-5,5), cex=24)
abline(v=1, lwd=2, col="red", add=T)
```

## Looking under the hood, DID

```{r underhood-did, height=6}
sim.out <- diagnosis$simulations
hist(sim.out$estimate[sim.out$estimator_label=="DID"],
     main="Randomization Distribution",
     xlab="Estimates in Realized Experiments",
     xlim=c(-3,3), cex=24)
abline(v=1, lwd=2, col="red")

#PH: note that we get more power using the difference in differences estimation. This is because diff-in-diff increases the precision of our estimate of the ATE.
```


